---
title: "Rock Paper Scissors"
output: html_notebook
---

Played Rock Paper Scissors at Evelyn's house

* I beat five or seven players in a row using dad's algorithm
* Is dad's algorithm actually better than randomly picking r p s?

```{r}
# library(tidyverse)
# looks like readr is broken at the moment
# see https://community.rstudio.com/t/issue-with-installing-the-drc-package/18701/3

```



```{r}
library(tidyr)
library(dplyr)
library(magrittr)
library(ggplot2)
```

# 1


Rules for our simulation

* Alice is our stand-in: every win, she gets a point, every tie she gets no points, and every loss she looses a point
* Bob is our opponent



Some simulation parameters and helper functions 

```{r}
nm <- c("rock", "paper", "scissors")
rps <- factor(nm, levels = nm)
n.games <- 60000
```


```{r}
# helper function, given vectors of alice and bob results, column binds and calculating the score of each game 
calc.wins <- function (alice, bob) {
    games <- bind_cols(list(alice, bob)) %>% 
        rename(alice = V1, bob = V2) %>% 
        mutate(score = case_when(
            alice == "rock" & bob == "rock" ~ 0
            , alice == "paper" & bob == "rock" ~ 1
            , alice == "scissors" & bob == "rock" ~ -1
            , alice == "rock" & bob == "paper" ~ -1
            , alice == "paper" & bob == "paper" ~ 0
            , alice == "scissors" & bob == "paper" ~ 1
            , alice == "rock" & bob == "scissors" ~ 1
            , alice == "paper" & bob == "scissors" ~ -1
            , alice == "scissors" & bob == "scissors" ~ 0
            , TRUE ~ 0
        ))
    return(games)
}
```


# 2

In a set of purely random games, we would not expect a clear winner or looser, e.g. uniform distribution of -1, 0, 1

so we simulate the games

```{r}
alice <- sample(rps, n.games, replace = TRUE)
bob <- sample(rps, n.games, replace = TRUE)
games <- calc.wins(alice, bob)
```

So let's see what the distribution of the scores were from the simulation.

```{r}
games %>% count(score)
```

```{r}
games %>% 
    ggplot(aes(x = factor(score))) +
    geom_bar() + 
    labs(title = "Looks like a uniform distribution on the scores", x = "Score", y = "Count", caption = "from random vs random")
```

Well certainly looks uniform, but how to be more precise about it? Let's use a chi-squared test. 

In a chi-squared test, we're testing the chances, under the null hypothesis that we should have had a uniform distribution, that we'd see this distribution by random chance

```{r}
games %>% 
    count(score) %>% 
    spread(key = score, value = n, fill = 0) %>% 
    chisq.test()
```

ok, so the p-value is very high, greater than 70% of the time, we'd see a distribution like this just by chance.


... what if we did this using resampling technique? Actually this breaks RStudio cloud

```{r}
# games %>%
#     count(score) %>%
#     spread(key = score, value = n, fill = 0) %>%
#     chisq.test(simulate.p.value = TRUE)
```


... what if we did this test by hand? 

Well the Pearson residuals for each cell is: (actual - expected) / sqrt(expected)

```{r}
games %>% 
    count(score) %>% 
    spread(key = score, value = n, fill = 0) %>% 
    mutate(
        `-1` = (`-1` - n.games/3) / sqrt(n.games/3)
        , `0` = (`0` - n.games/3) / sqrt(n.games/3)
        , `1` = (`1` - n.games/3) / sqrt(n.games/3))
```

And the chi-squared statistic is defined as the sum of the squared Pearson residuals

```{r}
games %>% 
    count(score) %>% 
    spread(key = score, value = n, fill = 0) %>% 
    mutate(
        `-1` = (`-1` - n.games/3) / sqrt(n.games/3)
        , `0` = (`0` - n.games/3) / sqrt(n.games/3)
        , `1` = (`1` - n.games/3) / sqrt(n.games/3)) %>% 
    gather(key = score, value = residuals) %>% 
    summarize(c = sum(residuals))
```




# 3

But what if Alice uses the following algorithm?

* in first game, randomly pick

```{r}

```

